---
title: "NJ MSA"
output: html_document
date: "2024-05-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load_packages, message=FALSE, warning=FALSE, results = "hide"}
library(tidyverse)
library(sf)
library(raster)
library(knitr)
library(kableExtra)
library(tidycensus)
library(tigris)
library(FNN)
#library(QuantPsyc) # JE Note: in R 4.1, QuantPsyc package not available.
library(caret)
library(yardstick)
library(pscl)
library(plotROC) 
library(ggrepel)
library(pROC)
library(grid)
library(gridExtra)
library(viridis)
library(igraph)

plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  #panel.border=element_rect(colour="#F0F0F0"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.75),
  axis.ticks=element_blank())

mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))

palette2 <- c("#41b6c4","#253494")
palette4 <- c("#a1dab4","#41b6c4","#2c7fb8","#253494")
palette5 <- c("#ffffcc","#a1dab4","#41b6c4","#2c7fb8","#253494")
palette10 <- c("#f7fcf0","#e0f3db","#ccebc5","#a8ddb5","#7bccc4",
               "#4eb3d3","#2b8cbe","#0868ac","#084081","#f7fcf0")
```

```{r, warning = FALSE, message = FALSE}
#this function converts a column in to quintiles. It is used for mapping.
quintileBreaks <- function(df,variable) {
    as.character(quantile(df[[variable]],
                          c(.01,.2,.4,.6,.8),na.rm=T))
}

#This function can be used to convert a polygon sf to centroids xy coords.
xyC <- function(aPolygonSF) {
  as.data.frame(
    cbind(x=st_coordinates(st_centroid(aPolygonSF))[,1],
          y=st_coordinates(st_centroid(aPolygonSF))[,2]))
} 

#this function convert a raster to a data frame so it can be plotted in ggplot
rast <- function(inRaster) {
  data.frame(
    xyFromCell(inRaster, 1:ncell(inRaster)), 
    value = getValues(inRaster)) }
```


# Setup and Introduction
Delaware Valley - Greater Philadelphia or the Philadelphia Metropolitan Area
Has a core metropolitan area of 6.3 million residents in 2020. Delaware Valley is the eighth largest metropolitan regiona in the nation. 

For the purpose of this study, we are only focusing on the counties located withing Pennslyvania State. These counties are the following:

- Bucks County
- Chester County
- Delaware County
- Montgomery County
- Philadelphia County

```{r load_data, warning = FALSE, message = FALSE, results = "hide"}
install.packages("readxl")
library(readxl)

tracts22 <- st_read("https://drive.google.com/open?id=1AsbWiYck9WOlU4QKkV1Uu2g7B5PIQO-_&usp=drive_fs") 

lc_2011 = raster("\Users\luyiiwong\Documents\GitHub\LandUseModeling_HW5\Data\lc_2011_Resample.tif")
lc_2021 = raster("\Users\luyiiwong\Documents\GitHub\LandUseModeling_HW5\Data\lc_2021_Resample.tif")

lc_change <- lc_2011+lc_2021

```

## calculating land cover change
```{r}
reclassMatrix <- 
  matrix(c(
    0,12,0,
    12,24,1,
    24,Inf,0),
  ncol=3, byrow=T)
```

```{r, warning = FALSE, message = FALSE}
developed_2011 <- 
  reclassify(lc_2011,reclassMatrix)

developed_2021 <- 
  reclassify(lc_2021,reclassMatrix)

```

```{r, warning = FALSE, message = FALSE}

development_change <- developed_2011+developed_2021

hist(development_change)
```

```{r, warning = FALSE, message = FALSE}
development_change[development_change != 1] <- NA

ggplot() +
  geom_sf(data=njMSA) +
  geom_raster(data=rast(development_change) %>% na.omit, 
              aes(x,y,fill=as.factor(value))) +
  scale_fill_viridis(discrete=TRUE, name ="Land Cover\nChange") + 
  labs(title="Development land use change") +
  mapTheme
```

```{r, warning = FALSE, message = FALSE}
lc_change2 <- 
  reclassify(lc_change,reclassMatrix)

lc_change2[lc_change2 < 1] <- NA

names(lc_change2) <- "lc_change"

ggplot() +
  geom_sf(data=njMSA) +
  geom_raster(data=rast(lc_change2) %>% na.omit, 
              aes(x,y,fill=as.factor(value))) +
  scale_fill_viridis(discrete=TRUE, name ="Land Cover\nChange") + 
  labs(title="Development Land Use Change") +
  mapTheme
```

# creating the fishnet
```{r, warning = FALSE, message = FALSE}
njMSA_fishnet <- 
  st_make_grid(njMSA, 2000) %>%
  st_sf()

njMSA_fishnet <-
  njMSA_fishnet[njMSA,]
```
```{r, warning = FALSE, message= FALSE}
ggplot() +
  geom_sf(data=njMSA_fishnet) +
  labs(title="Fishnet, 2000 Foot Resolution") +
  mapTheme
```

```{r, warning = FALSE, message = FALSE}
#original version
changePoints <-
  rasterToPoints(lc_change2) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(njMSA_fishnet))

fishnet <- 
  aggregate(changePoints, njMSA_fishnet, sum) %>%
  mutate(lc_change = ifelse(is.na(lc_change),0,1),
         lc_change = as.factor(lc_change))

ggplot() +
  geom_sf(data=njMSA) +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)$x, y=xyC(fishnet)$y, colour=lc_change)) +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name = "") +
  labs(title = "Land Cover Development Change", subtitle = "As fishnet centroids") +
  mapTheme
```

## 2.3. Land Cover in 2011

```{r, warning = FALSE, message = FALSE}
ggplot() +
  geom_sf(data=njMSA) +
  geom_raster(data=rast(lc_2011) %>% na.omit %>% filter(value > 0), 
              aes(x,y,fill=as.factor(value))) +
  scale_fill_viridis(discrete=TRUE, name ="") +
  labs(title = "Land Cover, 2011") +
  mapTheme +
  theme(legend.direction="horizontal")
```

The table below shows the approach taken to recoded existing land cover codes into the categories used in our analysis. In the code block below new rasters are generated and `names` are applied. Naming ensures that when the raster is integrated with the fishnet, the column reflects the appropriate raster.

| Old_Classification             | New_Classification                                  |
|--------------------------------|-----------------------------------------------------|
| Open Space as well as Low, Medium and High Intensity Development | Developed |
| Deciduous, Evergreen, and Mixed Forest |  Forest |
| Pasture\Hay and Cultivated Crops | Farm |
| Woody and Emergent Herbaceous Wetlands | Woodlands |
| Barren Land, Dwarf Scrub, and Grassland\Herbaceous | Other Undeveloped |
| Water | Water |

```{r, warning = FALSE, message = FALSE}
developed <- lc_2011 == 21 | lc_2011 == 22 | lc_2011 == 23 | lc_2011 == 24
forest <- lc_2011 == 41 | lc_2011 == 42 | lc_2011 == 43 
farm <- lc_2011 == 81 | lc_2011 == 82 
wetlands <- lc_2011 == 90 | lc_2011 == 95 
otherUndeveloped <- lc_2011 == 52 | lc_2011 == 71 | lc_2011 == 31 
water <- lc_2011 == 11

names(developed) <- "developed"
names(forest) <- "forest"
names(farm) <- "farm"
names(wetlands) <- "wetlands"
names(otherUndeveloped) <- "otherUndeveloped"
names(water) <- "water"
```

```{r, warning = FALSE, message = FALSE}
aggregateRaster <- function(inputRasterList, theFishnet) {
  #create an empty fishnet with the same dimensions as the input fishnet
  theseFishnets <- theFishnet %>% dplyr::select()
  #for each raster in the raster list
  for (i in inputRasterList) {
  #create a variable name corresponding to the ith raster
  varName <- names(i)
  #convert raster to points as an sf
    thesePoints <-
      rasterToPoints(i) %>%
      as.data.frame() %>%
      st_as_sf(coords = c("x", "y"), crs = st_crs(theFishnet)) %>%
      filter(.[[1]] == 1)
  #aggregate to the fishnet
    thisFishnet <-
      aggregate(thesePoints, theFishnet, length) %>%
      mutate(!!varName := ifelse(is.na(.[[1]]),0,1))
  #add to the larger fishnet
    theseFishnets <- cbind(theseFishnets,thisFishnet)
  }
  #output all aggregates as one large fishnet
   return(theseFishnets)
  }
```

```{r, warning = FALSE, message = FALSE}
theRasterList <- c(developed,forest,farm,wetlands,otherUndeveloped,water)

aggregatedRasters <-
  aggregateRaster(theRasterList, njMSA_fishnet) %>%
  dplyr::select(developed,forest,farm,wetlands,otherUndeveloped,water) %>%
  mutate_if(is.numeric,as.factor)

aggregatedRasters %>%
  gather(var,value,developed:water) %>%
  st_cast("POLYGON") %>%    #just to make sure no weird geometries slipped in
  mutate(X = xyC(.)$x,
         Y = xyC(.)$y) %>%
  ggplot() +
    geom_sf(data=njMSA) +
    geom_point(aes(X,Y, colour=as.factor(value))) +
    facet_wrap(~var) +
    scale_colour_manual(values = palette2,
                        labels=c("Other","Land Cover"),
                        name = "") +
    labs(title = "Land Cover Types, 2011",
         subtitle = "As fishnet centroids") +
   mapTheme
```

## 2.4. Census Data
```{r}
# setting up api key 
census_api_key("b83a23afee4a8ed0fa131e449869e6577b87151e", overwrite = TRUE, install = TRUE)
```
- Bucks County
- Chester County
- Delaware County
- Montgomery County
- Philadelphia County

```{r}
# creating list for selected counties
selected_counties <- c("Bergen County, New Jersey", "Hudson County, New Jersey", "Passaic County, New Jersey" , 
                       "Middlesex County, New Jersey", "Monmouth County, New Jersey", "Ocean County, New Jersey", 
                       "Somerset County, New Jersey", "Essex County, New Jersey", "Union County, New Jersey",
                       "Morris County, New Jersey", "Sussex County, New Jersey", "Hunterdon County, New Jersey",
                       "Mercer County, New Jersey", "Warren County, New Jersey")

# pulling data from census for 2011
njpop_2011 <- get_acs(geography = "tract", 
          variables = c("B01003_001E"), 
          year = 2011, 
          state = "NJ", 
          county = c("Bergen", "Hudson", "Passaic", "Middlesex", "Monmouth", "Ocean", "Somerset",
                     "Essex", "Union", "Morris", "Sussex", "Hunterdon", "Mercer", "Warren"),
          geometry = TRUE, 
          output = "wide")  %>%
  rename(pop_2011 = B01003_001E) %>%
  dplyr::select(GEOID, NAME, pop_2011, geometry) %>%
  st_transform(crs = 3424)


# pulling data from census for 2021
njpop_2021 <- get_acs(geography = "tract", 
          variables = c("B01003_001E"), 
          year = 2021, 
          state = "NJ", 
          county = c("Bergen", "Hudson", "Passaic", "Middlesex", "Monmouth", "Ocean", "Somerset",
                     "Essex", "Union", "Morris", "Sussex", "Hunterdon", "Mercer", "Warren"),
          geometry = TRUE, 
          output = "wide") %>%
  rename(pop_2021 = B01003_001E) %>%
  dplyr::select(GEOID, NAME, pop_2021, geometry) %>%
  st_transform(crs = 3424)

## grid arrange tract 2011 v 2021
grid.arrange(
ggplot() +
  geom_sf(data = njpop_2011, aes(fill=factor(ntile(pop_2011,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=quintileBreaks(njpop_2011,"pop_2011"),
                   name="Quintile\nBreaks") +
  labs(title="Population, New Jersey (by tract): 2011") +
  mapTheme,

ggplot() +
  geom_sf(data = njpop_2021, aes(fill=factor(ntile(pop_2021,5))), colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=quintileBreaks(njpop_2021,"pop_2021"),
                   name="Quintile\nBreaks") +
  labs(title="Population, New Jersey (by tract): 2021") +
  mapTheme, ncol=2)

```

## interpolating population and fishnet
```{r, warning = FALSE, message = FALSE}
njMSA_fishnet <-
  njMSA_fishnet %>%
  rownames_to_column("fishnetID") %>% 
  mutate(fishnetID = as.numeric(fishnetID)) %>%
  dplyr::select(fishnetID)

fishnetPopulation11 <-
  st_interpolate_aw(njpop_2011["pop_2011"], njMSA_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(njMSA_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(pop_2011 = replace_na(pop_2011,0)) %>%
  dplyr::select(pop_2011)

fishnetPopulation21 <-
  st_interpolate_aw(njpop_2021["pop_2021"],njMSA_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(njMSA_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(pop_2021 = replace_na(pop_2021,0)) %>%
  dplyr::select(pop_2021)

fishnetPopulation <- 
  cbind(fishnetPopulation11,fishnetPopulation21) %>%
  dplyr::select(pop_2011,pop_2021) %>%
  mutate(pop_Change = pop_2021 - pop_2011)
```

```{r, warning = FALSE, message = FALSE, fig.height = 8, fig.width= 11}
grid.arrange(
ggplot() +
  geom_sf(data=njpop_2021, aes(fill=factor(ntile(pop_2021,5))),colour=NA) +
  scale_fill_manual(values = palette5,
                    labels=substr(quintileBreaks(njpop_2021,"pop_2021"),1,4),
                   name="Quintile\nBreaks") +
  labs(title="Population, NJ MSA: 2021",
       subtitle="Represented as tracts; Boundaries omitted") +
  mapTheme,

ggplot() +
  geom_sf(data=fishnetPopulation, aes(fill=factor(ntile(pop_2021,5))),colour=NA) +
  scale_fill_manual(values = palette5,
                   labels=substr(quintileBreaks(fishnetPopulation,"pop_2021"),1,4),
                   name="Quintile\nBreaks") +
  labs(title="Population, NJ MSA: 2021",
       subtitle="Represented as fishnet gridcells; Boundaries omitted") +
  mapTheme, ncol=2)
```

# Highway Distance
```{r, warning = FALSE, message = FALSE, results = "hide"}
#njHighways <-
#  st_read("C:\Users\ferna\OneDrive\Documents\ArcGIS\Projects\CPLN 6750\HW5\dv_roads.geojson") %>%
#  st_transform(st_crs(njMSA)) %>%
#  st_intersection(njMSA)

njHighways <-
  st_read("\Users\luyiiwong\Documents\GitHub\LandUseModeling_HW5\Data\NEED_njroads_MSA.geojson") %>%
  st_transform(st_crs(njMSA)) %>%
  st_intersection(njMSA)
```

```{r plot_highway, warning = FALSE, message= FALSE}
ggplot() +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)[,1], y=xyC(fishnet)[,2],colour=lc_change),size=1.5) +
  geom_sf(data=njHighways) +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","New Development")) +
  labs(title = "New Development and Highways",
       subtitle = "As fishnet centroids") +
  mapTheme
```

```{r, warning = FALSE, message = FALSE}
emptyRaster <- lc_change
emptyRaster[] <- NA

njHighways_spdf <- as(njHighways, "Spatial")
highway_raster <- rasterize(njHighways, emptyRaster)

#highway_raster <- 
  #as(njHighways,'Spatial') %>%
  #rasterize(.,emptyRaster)

highway_raster_distance <- distance(highway_raster)
names(highway_raster_distance) <- "distance_highways"

highwayPoints <-
  rasterToPoints(highway_raster_distance) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(njMSA_fishnet))

highwayPoints_fishnet <- 
  aggregate(highwayPoints, njMSA_fishnet, mean) %>%
  mutate(distance_highways = ifelse(is.na(distance_highways),0,distance_highways))

ggplot() +
  geom_sf(data=njMSA) +
  geom_point(data=highwayPoints_fishnet, aes(x=xyC(highwayPoints_fishnet)[,1], 
                                             y=xyC(highwayPoints_fishnet)[,2], 
                 colour=factor(ntile(distance_highways,5))),size=1.5) +
  scale_colour_manual(values = palette5,
                      labels=substr(quintileBreaks(highwayPoints_fishnet,"distance_highways"),1,8),
                      name="Quintile\nBreaks") +
  geom_sf(data=njHighways, colour = "red") +
  labs(title = "Distance to Highways",
       subtitle = "As fishnet centroids; Highways visualized in red") +
  mapTheme
```

#spatial lag
``` {r}
nn_function <- function(measureFrom,measureTo,k) {
  #convert the sf layers to matrices
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
    output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}
```

```{r}
fishnet$lagDevelopment <-
    nn_function(xyC(fishnet),
                xyC(filter(aggregatedRasters,developed==1)),
                2)

ggplot() +
  geom_sf(data=njMSA) +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)[,1], y=xyC(fishnet)[,2], 
                 colour= log(lagDevelopment), size=.001)) +
  labs(title = "Spatial Lag to 2011 Development",
       subtitle = "As fishnet centroids")
# make histogram
# try viridis
# log color\lagdevelopment
# filter out og developed cells 
```

# Create MSA Counties 

```{r, warning = FALSE, message = FALSE, results = "hide"}
options(tigris_class = "sf")

studyAreaCounties <- 
  counties("New Jersey") %>%
  st_transform(st_crs(njMSA)) %>%
  dplyr::select(NAME) %>%
  .[st_buffer(njMSA,-1000), , op=st_intersects] 

studyAreaCounties <- studyAreaCounties[studyAreaCounties$NAME != 'Atlantic', ]
```

```{r, warning = FALSE, message = FALSE}
ggplot() +
  geom_sf(data=studyAreaCounties) +
  labs(title = "Study Area Counties") +
  mapTheme
```
#  Create the Final Dataset 
Once we join the data set with out county boundaries, all the lc_change =1 is dropped
```{r}
dat <- 
  cbind(fishnet, highwayPoints_fishnet, fishnetPopulation, aggregatedRasters)%>%
  dplyr::select(lc_change, developed, forest, farm, wetlands, otherUndeveloped, water,
                pop_2011, pop_2021, pop_Change, distance_highways,lagDevelopment) %>%
  st_join(studyAreaCounties) %>%
  mutate(developed10 = ifelse(lc_change == 1 & developed == 1, 0, developed)) %>%
  filter(water == 0) 


dat <- cbind(fishnet, aggregatedRasters, st_drop_geometry(highwayPoints_fishnet), st_drop_geometry(fishnetPopulation), st_drop_geometry(fishnet)) %>%
  filter(water == 0)

tm_shape(dat) +
  tm_fill(col = "lc_change", palette = palette2, style = "cat")
```

# Exploratroy Analysis
It seems like land change only occured from undeveloped to water. Therefore, there is no new development...
```{r, warning = FALSE, message = FALSE}
dat %>%
  dplyr::select(distance_highways,lagDevelopment,lc_change) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
    facet_wrap(~Variable) +
    scale_fill_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New Development as a Function of the Continuous Variables") +
    plotTheme 
```


```{r, warning = FALSE, message = FALSE}
dat %>%
  dplyr::select(pop_2011,pop_2021,pop_Change,lc_change) %>%
  gather(Variable, Value, -lc_change, -geometry) %>%
  ggplot(., aes(lc_change, Value, fill=lc_change)) + 
    geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
    facet_wrap(~Variable) +
    scale_fill_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New Development as a Function of Factor Variables") +
    plotTheme
```

#### OLD Highway Distance OLD
```{r, warning = FALSE, message = FALSE, results = "hide"}
njHighways <-
  st_read("\Users\luyiiwong\Documents\Land Use & Environmental Modeling\Assignment5\dvroad_clip\dvroad_clip2.shp") %>%
  st_transform(st_crs(njMSA)) %>%
  st_intersection(njMSA)
```

```{r plot_highway, warning = FALSE, message= FALSE}
ggplot() +
  geom_point(data=fishnet, 
             aes(x=xyC(fishnet)[,1], y=xyC(fishnet)[,2],colour=lc_change),size=1.5) +
  geom_sf(data=njHighways) +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","New Development")) +
  labs(title = "New Development and Highways",
       subtitle = "As fishnet centroids") +
  mapTheme
```

```{r, warning = FALSE, message = FALSE}
emptyRaster <- lc_change
emptyRaster[] <- NA

njHighways_spdf <- as(njHighways, "Spatial")
highway_raster <- rasterize(njHighways, emptyRaster)

highway_raster <- raster(njHighways)

highway_raster <- 
  as(njHighways,'Spatial') %>%
  rasterize(.,emptyRaster)

highway_raster_distance <- distance(highway_raster)
names(highway_raster_distance) <- "distance_highways"

highwayPoints <-
  rasterToPoints(highway_raster_distance) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(houstonMSA_fishnet))

highwayPoints_fishnet <- 
  aggregate(highwayPoints, houstonMSA_fishnet, mean) %>%
  mutate(distance_highways = ifelse(is.na(distance_highways),0,distance_highways))

ggplot() +
  geom_sf(data=houstonMSA) +
  geom_point(data=highwayPoints_fishnet, aes(x=xyC(highwayPoints_fishnet)[,1], 
                                             y=xyC(highwayPoints_fishnet)[,2], 
                 colour=factor(ntile(distance_highways,5))),size=1.5) +
  scale_colour_manual(values = palette5,
                      labels=substr(quintileBreaks(highwayPoints_fishnet,"distance_highways"),1,8),
                      name="Quintile\nBreaks") +
  geom_sf(data=houstonHighways, colour = "red") +
  labs(title = "Distance to Highways",
       subtitle = "As fishnet centroids; Highways visualized in red") +
  mapTheme
```